---
title: "Food Insecurity using Household Living Budget - HLB"
author: "cm"
date: "09/24/2023" #update from the previous code
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Start fresh
```{r}
rm(list=ls())
```

# Libraries
```{r}
#install.packages("ipumsr")
library(ipumsr)
#install.packages("tidyverse")
#library(tidyverse)
library(survival)
library(ggfortify)
library(dplyr)
library(readxl)
library(readr)
library(reshape2)
library(tidyr)
```

#  ######################## 1 DETAILS: Specify the fips code for the State and county ########################
```{r}

county_fips <- 11001  #District of Columbia - Washington DC
county_name <- "DC"
name_files <- "dc"
state_name <- "dc"

state_fip <- substr(county_fips, 1,2)
county_fip <- substr(county_fips, 3,5)

```

## Data
```{r}
######## direct upload with repo
#data_syn_pop <- read_csv("~/Documents/food_insecurity/source_code/analyses/dc_pums_synpop_2021.csv.zip")

######## upload from url
#url-raw info
url_synpop <- "https://github.com/uva-bi-sdad/food_insecurity/raw/main/documents/products/data_tables/dc_pums_synpop_2021.csv.zip"

#destination file: CHOOSE anywhere you want to download it
destfile_synpop <- "~/Documents/food_insecurity/documents/products/derived_variables/download_dc_2021.csv.zip"

#download file and save in specified destination
download.file(url_synpop, destfile_synpop)

#open csv zip
data_syn_pop <- read_csv("~/Documents/food_insecurity/documents/products/derived_variables/download_dc_2021.csv.zip") 

#data set for hlb
dat_hlb <- data_syn_pop
 
#add row number as serial: this is the identifier for every household in the county
dat_hlb$serial <- seq.int(nrow(dat_hlb))
 
```

#  ######################## 2 Household Living Budget Components - HLB  ########################

# 2.1 Exercise at the household level

# Food Cost 
```{r}
#source:
#https://www.fns.usda.gov/cnpp/usda-food-plans-cost-food-reports-monthly-reports
#https://fns-prod.azureedge.us/sites/default/files/media/file/CostofFoodAug2022LowModLib.pdf
#adjustmen Feeding America
#https://map.feedingamerica.org/county/2020/overall/virginia

#data counties
food_cost_adj <- read_excel('~/Documents/food_insecurity/data/household_living_budget/food/va_ct_feeding_america_map_meal_gap_2022SEP.xlsx')

#data USDA USA
usda_meal_plan_sep2022 <- read_csv("~/Documents/food_insecurity/data/household_living_budget/food/national_usda_meal_plans_2022SEP.csv", skip = 1)

#data USDA categories and cost
usda_low_cost_meal <- usda_meal_plan_sep2022 %>% select(group=`...2`, monthly_cost=`Low-cost...6`) %>% filter( !is.na(monthly_cost)) %>% filter(monthly_cost!= 'plan')

#remove $
usda_low_cost_meal$monthly_cost <- gsub( '\\$' ,'', usda_low_cost_meal$monthly_cost )

#vector with costs
food_cost_general <-  data.frame(
no_adult=  mean( c(as.numeric( usda_low_cost_meal %>% filter(group=='19-50 years') %>% pull(monthly_cost) ) ) ),
no_infant= as.numeric( usda_low_cost_meal %>% filter(group=='1 year') %>% pull(monthly_cost) ) ,
no_toddler=  mean( c(as.numeric( usda_low_cost_meal %>% filter(group=='2-3 years') %>% pull(monthly_cost) ) ) ),
no_preschooler= mean(as.numeric(usda_low_cost_meal %>% filter(group=='4-5 years') %>% pull(monthly_cost)) ),
no_schooler= mean(c( as.numeric(usda_low_cost_meal %>% filter(group=='6-8 years') %>% pull(monthly_cost)) ,
                     as.numeric(usda_low_cost_meal %>% filter(group=='9-11 years') %>% pull(monthly_cost)) )),
no_teenager=  mean(c( as.numeric(usda_low_cost_meal %>% filter(group=='12-13 years') %>% pull(monthly_cost)) ,
                     as.numeric(usda_low_cost_meal %>% filter(group=='14-18 years') %>% pull(monthly_cost)) )) )


#ADJUSTMENTS
#AVERAGE MEAL COST IN THE UNITED STATES: $3.25. Source: Feeding America for 2020
meal_avg_cost_us <- 3.25
#cost for county
#https://map.feedingamerica.org/county/2020/overall/district-of-columbia reference
meal_avg_cost_county_va <- 4.16

#adjustment
adj_food_cost <- meal_avg_cost_county_va /meal_avg_cost_us
#food inflation - BLS to adjust meal cost from 2020
inflation_usa_food_2021 <- 0.065
inflation_usa_food_2022 <- 0.118
inflation_msa_mdv_food_2021 <- 0.057
inflation_msa_mdv_food_2022 <- 0.087

#adjustment + inflation_food adjustment
adj_food_cost_inf <- meal_avg_cost_county_va/meal_avg_cost_us*(1+inflation_msa_mdv_food_2021)*(1+inflation_msa_mdv_food_2022)/((1+inflation_usa_food_2021)*(1+inflation_usa_food_2022))

#food cost adjusted by geography (county) and by inflation
food_cost_general_adj_infl <- adj_food_cost_inf*food_cost_general


#vector of per person cost - food
vector_food_cost <- data.frame(cost_pc = t(food_cost_general_adj_infl))
vector_food_cost <- tibble::rownames_to_column(vector_food_cost, "person")

#food cost long format
food_cost_long <- dat_hlb  %>% select(puma, serial, hh_income_cat, hh_size_recode, no_adult, no_preschooler, no_toddler, no_schooler, no_teenager, no_infant ) %>% arrange(serial) %>% pivot_longer(cols=c('no_adult', 'no_preschooler', 'no_toddler', 'no_schooler', 'no_teenager', 'no_infant'),
                    names_to='person',
                    values_to='num_person') %>% left_join(vector_food_cost, by = "person") %>% mutate(adj_econ_scales = case_when( 
                      hh_size_recode == 1 ~ 1 + 0.2, 
                      hh_size_recode == 2 ~ 1 + 0.1,
                      hh_size_recode == 3 ~ 1 + 0.05,
                      hh_size_recode == 4 ~ 1 + 0,
                      hh_size_recode == 5 ~ 1 - 0.05)
                      ) %>% mutate(food_cost_month = num_person*cost_pc*adj_econ_scales)

#food cost
food_hlb <- food_cost_long %>% group_by(serial) %>% summarise(food_cost_month = sum(food_cost_month))

```

# Childcare Cost
```{r}
#data -  Dept of Labor
childcare_dol <- read_excel("~/Documents/food_insecurity/data/household_living_budget/childcare/dmv_dol_childcare_costs_2022.xlsx")
#data county
childcare_dol_county <- childcare_dol %>% filter(year==2022) %>% filter(FIPS == county_fips) %>%
  select("name_county", "FIPS", "year", "infant_family", "toddler_family", "preschooler_family", "schooler_family" )


#weekly cost
cc_cost_indiv_min <- c(
adult= 0, 
infant= childcare_dol_county %>% pull(infant_family),
toddler= childcare_dol_county %>% pull(toddler_family),
preschooler= childcare_dol_county %>% pull(preschooler_family),
schooler= childcare_dol_county %>% pull(schooler_family),
teenager=0) 


# cost of child care per representative family - min - per month
childcare_hlb <-  dat_hlb %>% mutate(childcare_cost_month= 
                                                         as.numeric(cc_cost_indiv_min['adult'])*no_adult*4  +
                                                         as.numeric(cc_cost_indiv_min['teenager'])*no_teenager*4 +
                                                         as.numeric(cc_cost_indiv_min['schooler'])*no_schooler*4 +
                                                         as.numeric(cc_cost_indiv_min['preschooler'])*no_preschooler*4 +
                                                         as.numeric(cc_cost_indiv_min['toddler'])*no_toddler*4 +
                                                         as.numeric(cc_cost_indiv_min['infant'])*no_infant*4
                                                         ) %>% select(serial, childcare_cost_month )

```

# Housing Cost 
```{r}
#data fro HUD
housing_cost_updated <- read_excel("~/Documents/food_insecurity/data/household_living_budget/housing/dc_tr_hud_2022_housing_cost.xlsx") 

housing_cost_updated <- housing_cost_updated %>% mutate(fips=substr(geoid,1,5)) %>% filter(fips %in% county_fips )

housing_cost_updated_wide <- spread(housing_cost_updated, key = measure, value = value)

housing_cost_county <- housing_cost_updated_wide %>% mutate(
  `1` =  monthly_rent_0br,
  `2` =  monthly_rent_1br,
  `3` =  monthly_rent_2br,
  `4` =  monthly_rent_3br,
  `5` =  monthly_rent_4br
  ) %>% select( -c(monthly_rent_0br,monthly_rent_1br,monthly_rent_2br,monthly_rent_3br,monthly_rent_4br, measure_type, region_type, region_name, year, fips) )


#housing cost in long format
housing_cost_long <- housing_cost_county %>% pivot_longer(cols=c(`1` ,  `2`  , `3` ,  `4` ,  `5`),
                    names_to='hh_size_recode',
                    values_to='housing_cost_month') %>% mutate(tract= substr(geoid, 6, 11))

#join to match housing by place (tract) and household size
dat_hlb$tract <- as.character(dat_hlb$tract)
dat_hlb$hh_size_recode <- as.character(dat_hlb$hh_size_recode)

#food cost per household (serial)
housing_hlb <- dat_hlb %>% left_join(housing_cost_long, by = c("tract", "hh_size_recode")) %>% select(serial, housing_cost_month)

```

# Cost of Transportation
```{r}
#original data from CNT

htaindex2020_data_tracts <- read_csv("~/Documents/food_insecurity/data/household_living_budget/transportation/htaindex2020_data_tracts_11.zip")

#remove " from geoid
htaindex2020_data_tracts$tract <- gsub( '\\"' ,'', htaindex2020_data_tracts$tract )


#create fips code
htaindex2020_data_tracts$fips <- substr(htaindex2020_data_tracts$tract, 1, 5)
#filter by county and relevant data (data by year)
transp_cost_county <- htaindex2020_data_tracts %>% filter(fips == county_fips) %>% select(geoid = tract, t_cost_ami)
#cost estimate per person: National Typical Household: 
#we are assuming 2.62 as average size in order to account for individual cost
#adjustment to 2022: https://data.bts.gov/stories/s/Transportation-and-Inflation/f9jm-cqwe#
cpi_transpor2020 <- 112.9 #june
cpi_transpor2022 <- 134.9 #june 
transp_cost_county <- transp_cost_county %>% mutate(transp_cost_person = t_cost_ami/2.72/12*cpi_transpor2022/cpi_transpor2020, 
                                                    tract = substr(geoid, 6, 11) )

#total estimation
transportation_hlb_init <- dat_hlb %>% left_join(transp_cost_county, by = c("tract")) %>% mutate(transp_cost_month = transp_cost_person*no_adult ) 

#transportation cost for every household
transportation_hlb <- transportation_hlb_init %>% select(serial, transp_cost_month)

#lab
#transportation_hlb_na <- transportation_hlb_init %>% filter(is.na(trasp_cost_mont))

```
#Broadband cost 
```{r}
#data
dat_broadband <- read.csv("~/Documents/food_insecurity/data/household_living_budget/broadband/dcmdva_hdcttrbg_2021_broadband_now_internet_package_price.csv.xz")

#county id
dat_broadband$county <- substr(dat_broadband$geoid,1,5)

library(dplyr)
#use min price 25
bband_county_tr <- dat_broadband %>% filter(county %in% county_fips) %>% filter(region_type =='tract') %>% filter(measure=="min_price_25")

#price increase escenario
price_increase_wireless_service <- 0  #0.013

#price per tract
bband_tract <- bband_county_tr %>% mutate(broadband_cost_month = value*(1+price_increase_wireless_service)) %>% select(geoid, broadband_cost_month) %>% mutate(tract= substr(geoid, 6, 11))

#prices all tracts
tracts_county <- data.frame( tract = unique(dat_hlb$tract) )
bband_tract_prices <- tracts_county %>% left_join(bband_tract, by = "tract")

#fill NAs with value close to the tract
#library(tidyr)
bband_tract_prices <- fill(bband_tract_prices, broadband_cost_month, .direction = 'up')
bband_tract_prices <- bband_tract_prices %>% select(tract, broadband_cost_month)
#estimation
broadband_hlb_init <- dat_hlb %>% left_join(bband_tract_prices, by = "tract")

#broadband cost for every household
broadband_hlb <- broadband_hlb_init %>% select(serial, broadband_cost_month)

```


#Merge with df_hlb
```{r}

#food cost
df_hlb6 <-  dat_hlb %>% left_join(food_hlb %>% select(serial, food_cost_month) , by = 'serial')

#childcare
df_hlb6 <- df_hlb6 %>% left_join( as.data.frame(childcare_hlb) %>% select(serial, childcare_cost_month) , by = 'serial') 

#housing
df_hlb6 <- df_hlb6 %>% left_join( housing_hlb , by = 'serial')

#transportation
df_hlb6 <- df_hlb6 %>% left_join( transportation_hlb , by = 'serial')

#broadband
df_hlb6 <- df_hlb6 %>% left_join( broadband_hlb , by = 'serial')

#other costs
percentage_other_expenses <- 0.2
df_hlb6 <- df_hlb6 %>% mutate(other_costs_month = (food_cost_month+housing_cost_month)*percentage_other_expenses ) 

#formats
df_hlb6$serialno <- as.factor(df_hlb6$serialno)
df_hlb6$hh_comb <- as.factor(df_hlb6$hh_comb)
```


#  ######################## 3 Data set  ########################

# Save data set of HLB with information of census tracts, income category and household combinations
# With this file, proceed to sample to create the synthetic population
```{r}

#save file in 
if(!file.exists( paste0("~/Documents/food_insecurity/documents/products/derived_variables/", name_files, "_hlb6_2021.csv") )){
  
  write.csv(df_hlb6, paste0("~/Documents/food_insecurity/documents/products/derived_variables/", name_files, "_hlb6_2021.csv") , row.names = FALSE)
}

```
